\lecture{Nov. 21}

\begin{defn}[RAS]
    Key Generation: Randomly pick $p$,$q$ primes $n = pq$ pick $e \cdot d = 1 \mod \phi (n)$
    Encryption: $c = m^e \mod \phi (n)$
\end{defn}

\topic{Some attack on RSA}
Collect a lot of $n_i = p_i\cdot q_i$. Compute gcd of $n_i,j_i$ where $i\neq j$. Some gcds are not equal to $1$ and thus $n_i$ can be factored.


\begin{align*}
    &\text{number of primes} < 2^{512} \approxeq \frac{2^{512}}{512\cdot log2}\\
    &\text{number of primes} < 2^{511} \approxeq \frac{2^{511}}{511 \cdot log2}\\
    &\text{number of primes with 512 bits} \approxeq {2^{500}}
\end{align*}

\begin{exmp}
    Sometimes e = 3
    
    Advantage: faster encryption
    
    Disadvantage: $n \approxeq 2^{2048}$ if $m < 2^{600} m^3 \mod n = m^3$ as integer
    
    In practice: Padding of m is about 600, where the total from 1~random~m is about 2000. 
\end{exmp}


\topic{Digital Signature}
\begin{enumerate}
    \item Authentic
    \item Alice which is the sender cannot deny the message she sent (non-repudiation)
\end{enumerate}


\begin{exmp}[Naive TSA Signature]\leavevmode
(Where Alice sent a message to Bob and Eve is the outsider)

    (n,e) is a public key for Alice
    
    d is a private key for Alice
    
    \[S = m^d \mod n\]
    
    Bob will verify by comparing $S^e \mod n$ (where $S$ is the signature).
\end{exmp}


\topic{Attack Models}
\begin{enumerate}
\item Key-Only Attack: Eve only knows Alice's public key
\item Known-Message attack: Eve knows some $(m_i,s_i)$
\item Chosen-message attack (CMA): Eve can obtain signature $s_i$ for arbitrary message $m_i$.
\item Totally broken: Eve can sign any message $m$.
\item Selection Forgery: Eve can sign one message of her choice.
\item Existential Forgery (ET): There exist a message that Eve can sign.

\begin{note}
    We call Digital Signature a secure if Eve cannot achieve ET using CMA.
\end{note}
\end{enumerate}

\begin{claim}
    For the pervious exmaple:
    (1,1) is always valid, and thus we claim that it is totally broken under CMA
\end{claim}

\begin{proof}
    Given any $m$, Pick $a, b\neq 1$ such that $a\cdot b = m \mod n$
    
    Eve can obtain $s_1 = a^d \mod n$ and $s_2 = b^d \mod n$. Then $s_1\cdot s_2 = (ab)^d = m^d \mod n$.
\end{proof}

To make it more secure, we will apply some functions on the message on called the hash function.
\begin{exmp}[Hash Function]
    $H: \{0, 1\}^k \to \{0, 1\}^n$ is takes an infinite set to a finite set.

    Preimage Resistant: for every y, it is hard to find H(m)= y
    
    2nd Preimage Resistant: for every value of $m$, it is hard to find $m' \neq m$ such that $H(m) = H(m')$
    
    Collision Resistant: it is hard to hard $m$ and $m'$ with $H(m) = H(m')$

    Note: Collision Resistant implies 2nd Preimage Resistant.
    For such function, it should occur that $H(a,b) \neq H(a)\cdot H(b)$
\end{exmp}


\begin{exmp}
    if H is not preimage resistant, then Eve can find m such that H(m) = 1
    
    Since 1 is a signature for m, if H is not collision resistant, Eve can compute m, m' a collision
    
    Under a CMA, request signature for m' that's also signature for m.
\end{exmp}