\lecture{Nov. 14}


\begin{defn}
    Assume that $f(x)$ is differentiable at each $x_0$ in an interval I. We define $f': I \to R$ by \[f'(x)= \lim_{h \to 0} \frac{f(x_0+h)-f(x_0)}{h}\]. $f'$ is called the derivative (function) of f on I.
\end{defn}

\begin{exmp}
    $f(x) = sinx$, $f'(x) = cosx$ on $\mathbb{R}$
\end{exmp}


\begin{nota}\leavevmode

\begin{enumerate}
    \item $y=f(x) \to y'$ will denote $f'(x)$
    \item $\displaystyle \frac{dy}{dx} = f'(x)$
    \item $\displaystyle \frac{d}{dy} f(x) = f'(x)$
\end{enumerate}
\end{nota}

If $f'(x)$ is differentiable at $x_0\in \mathbb{R}$, then we call \[(f')'(x_0) = \lim_{h\to 0}\frac{f'(x_0+h) - f'(h)}{h}\] the second derivative of $f$ at $x=x_0$. We denote this by $f''(x_0)$.

In general if $f$ is twice differentiable on $I$, we write $f''(x)$ to represent the second derivative.

$f'''(x) \to $ third derivative.

$f^{(n)}(x)$ denotes the n-th derivative.

\begin{thm}[More on Linear Approximation]
    If $f(x)$ is differentiable at $x=a$, and if \[L_a(x) = f(a) +f'(a)(x-a)\] then $L_a(x)\approxeq f(x)$ if $ x\approxeq a$
\end{thm}

\begin{thm}[Error in Linear Approximation]
    \[\text{Error} = \abs{f(x) - L_a(x)}\]
    
The error is effected by
\begin{enumerate}
\item Distance of $x$ to $a$.
\item The larger $\abs{f''(x)}$ the larger the error may be.
\end{enumerate}
Both 1 and 2 hold in general most of the time but not always.
\end{thm}

\begin{thm}[Newton's Method]
Pick $a_1$.

Let \[a_{n+1} = a_n - \frac{f(a_n)}{f'(a_n)}\]
\end{thm}

\begin{rem}\leavevmode

\begin{enumerate}
\item If $f'(c)\neq 0$, then there exists $\delta > 0$ such that if $a_1 \in (c-\delta,c+\delta)$ then $a_n \to c$
\item When the method work, the convergence is generally very fast. In general, the convergence is ``quadratic" in nature. Roughly speaking this means the number of decimal places of accuracy will at least double with each iteration.
\item It can fail.
\end{enumerate}
\end{rem}

\begin{exmp}[Heron's Method]
Solve $x^2 - a = 0$.
\end{exmp}

\begin{solution}
Pick $a_1$. \[a_{n+1} = a_{n} - \frac{f(a_n)}{f'(a_n)} = a_n - \frac{a^2_n - a}{2a_n} = \frac{1}{2} (a_n + \frac{a}{a_n})\]
\end{solution}